Fitting parametric univariate distributions to non censored or censored data using the R \texttt{fitdistrplus} package

### 2012-03-21 inst/doc/intro2fitdistrplus.Rnw
correction of the problem with optim using "L-BFGS-B" and new version of the vignette

Fitting distributions to data is a very common task in statistics. It consists of choosing
a probability distribution that gives a good representation of a statistical
variable. It requires judgment and expertise and generally needs an iterative process of
distribution choice, parameter estimation, and quality of fit evaluation. 
Function \texttt{fitdistr} in the R package \texttt{MASS} \cite{MASS} 
is a well known general-purpose maximum-likelihood
fitting routine for the parameter estimation step in R. Other steps of
the process may be developed using R \cite{Ricci05}. Our first objective while developping
package \texttt{fitdistrplus} \cite{fitdistrplus} 
was to provide R users a set of functions dedicated to help the overall
process of fitting a univariate parametric distribution to data.

@@

% A changer en cas de mise ‡ jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
This paper reviews the various features of version 0.3-0 of \texttt{fitdistrplus}. The package
is available from the Comprehensive R Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
The development version of the package is located at R-forge as one the packages of the 
project ``Risk Assessment with R''
(\url{http://r-forge.r-project.org/projects/riskassessment/})
The following command will load the package.
<<results=hide>>=
library(fitdistrplus)
@

\subsection{Running examples}
\label{Examples}
%%%%%%%%%%%
For illustrating the use of various functions of package \texttt{fitdistrplus}, we will use
four examples published in various biological areas, 
corresponding to data sets included in the package.
% voir si on en met 3 ou 4 au final !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@

The third data set corresponds to the observation of a discrete variable, 
the number of \emph{Toxocara cati} parasites present in digestive tract,
on a random sample of feral 
cats living on Kerguelen island (\cite{Fromont01}).
<<echo=TRUE>>=
data(toxocara)
str(toxocara)
@

The last data set corresponds to the observation of a continuous censored variable, 
the \emph{Listeria monocytogenes} microbial concentration, on a random
sample of smoked fish distributed on the Belgian market in the period
2005 to 2007 (\cite{Busschaert10}).
Censored data are coded within 2 columns named left and right, describing
each observed value of \emph{Listeria monocytogenes} concentration 
(in $CFU.g^{-1}$) as an interval. 
The left column contains either \texttt{NA} for
left censored observations, the left bound of the interval for interval censored
observations, or the observed value for non-censored observations. The right
column contains either \texttt{NA} for right censored observations, the right bound of
the interval for interval censored observations, or the observed value for noncensored
observations.
<<echo=TRUE>>=
data(smokedfish)
str(smokedfish)
@

@@

<<echo=TRUE, fig=FALSE>>=
plotdist(groundbeef$serving)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, fig=TRUE>>=
plotdist(groundbeef$serving)
@
  \caption{Density and cdf plots of an empirical distribution for a continuous variable}
  \label{plotdistcont}
\end{figure}

@@

<<echo=TRUE, fig=FALSE>>=
plotdist(toxocara$number,discrete = TRUE)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, fig=TRUE>>=
plotdist(toxocara$number,discrete = TRUE)
@
  \caption{Density and cdf plots of an empirical distribution for a discrete variable}
  \label{plotdistcont}
\end{figure}

@@

\subsection{Empirical basis for selecting candidate distributions}
\label{Empirical}
%%%%%%%%%%
Function \texttt{descdist} 
Descriptives statistics may help the choice of good candidates to describe an empirical distribution
among a family of parametric distributions. Especially the skewness and kurtosis are useful for this
purpose. The concept of skewness relates to deviations from symmetry of the distribution.
The normal distribution has a skewness of zero. A positive (resp. negative) skewness indicates 
that the right (resp. left) tail of the distribution is more extended than the left (resp. right) one.
The concept of kurtosis relates to the tail weight. The normal distribution has a kurtosis of 3.
Distributions with a higher kurtosis are said to be leptokurtic, with heavier tails, such as the 
logistic distribution, while distributions with a smaller kurtosis are said platykurtic, with lighter 
tails, such as the uniform distribution.

Function \texttt{descdist} provides calculations of classical descriptive statistics
(minimum, maximum, median, mean, sample sd) and by default 
unbiased estimations of skewness and 
Pearsons's kurtosis values. Nevertheless estimations of skewness and kurtosis
are unbiased only for normal distributions and estimated values are thus only indicative.
A skewness-kurtosis plot such as the one proposed by \cite{Cullen99} is also provided for the 
empirical distribution (Figure~\ref{Cullenplotcont}). 
On this plot, values for common distributions are displayed as a tools 
to help the choice of distributions to fit to data. For some distributions (normal, uniform,
logistic, exponential for example), there is only one possible value for the skewness and the kurtosis
 and the distribution is thus represented by a point on the plot. For other distributions, 
areas of possible values are represented, consisting in lines (as for gamma and lognormal distributions), 
or larger areas (as for beta distribution).

Skewness and kurtosis are known not to be robust. In order to take into account the uncertainty 
of the estimated values of kurtosis and skewness from data, the data set may be boostraped by 
fixing the argument \texttt{boot} to an integer above 10. Values of skewness and kurtosis 
corresponding to bootstrap samples are then computed and reported on the 
skewness-kurtosis plot.

<<echo=TRUE, fig=FALSE>>=
descdist(groundbeef$serving,boot=1000)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
descdist(groundbeef$serving,boot=1000)
@
  \caption{Skewness-kurtosis plot for a continuous variable}
  \label{Cullenplotcont}
\end{figure}

For discrete variable, skewness and kurtosis values or set of values of Poisson and negative binomial
distributions are represented in the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}), 
together with values for the
normal distribution, to which discrete distributions may converge.

<<echo=TRUE, fig=FALSE>>=
descdist(toxocara$number,discrete = TRUE,boot=1000)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
descdist(toxocara$number,discrete = TRUE,boot=1000)
@
  \caption{Skewness-kurtosis plot for a continuous variable}
  \label{Cullenplotdisc}
\end{figure}

@@

\section{Fit of a distribution}
\label{FIT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter estimation}
\label{Estim}
%%%%%%%%%%
Once selected, one or more parametric distributions may be fitted to the data set, one at a time, using
Function \texttt{fitdist}. By default, distribution parameters $\theta$ are estimated by maximizing the
likelihood defined as:
  \begin{equation}
    \label{likelihood}
    L(\theta)=\prod_{i=1}^n f(y_{i}|\theta)
  \end{equation}
with $y_{i}$ the $n$ observations of variable $y$ and $f$ the density function of the fitted 
parametric distribution.

@@

Function \texttt{fitdist} returns the results of the fit of any parametric distribution 
to a data set as an S3 class object that may be easily printed, 
summarized or plotted (see Figure~\ref{fitgamma} in Section~\ref{GOF plots}). 
The parametric distribution must be a classically defined R distributions,
with at least d, p and q functions respectively for the density cdf and quantile functions
(for example \texttt{dnorm}, \texttt{pnorm} and \texttt{qnorm} for the normal distribution).
The name of the fitted distribution is specified in the first argument by its classical 
abbreviation used as the second part of d, p and q functions (for example ``norm''
for the normal distribution. Numerical results returned by Function \texttt{fitdist} are 
parameter estimates with estimated standard errors computed from the estimate
of the Hessian matrix at the maximum likelihood solution, correlation matrix
between parameter estimates, the loglikelihood, the AkaÔke and the Schwarz information criteria 
(so called AIC and BIC).
<<echo = TRUE,fig = FALSE>>=
fw <- fitdist(groundbeef$serving,"weibull")
print(fw)
summary(fw)
@

@@

\subsection{Goodness-of-fit plots}
\label{GOF plots}
%%%%%%%%%%
The plot of an object of class \texttt{fitdist} provides two types of results depending of the nature 
of the distribution, continuous or discrete. For continuous distributions, 
four goodness-of-fit plots are provided : a draw of pdf curve and histogram together, an cdf plot
of both empirical and theoretical distributions, a Q-Q plot
(plot of the quantiles of the theoretical fitted distribution (x-axis) against the empirical quantiles of the data)
and a P-P plot (i.e. for each value of the data set, plot of the cumulative density function of the fitted distribution
(x-axis) against the empirical cumulative density function (y-axis)) are also given (\cite{Cullen99}).
As an exemple, let us look at the plot of the previous fit of a weibull distribution to ``groundbeef'' data set.

<<echo=TRUE, fig=FALSE>>=
plot(fw)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(fw)
@
  \caption{Plot of the fit of a continuous distribution}
  \label{plotcontfit}
\end{figure}

For continuous distributions, Function \texttt{cdfcomp} enables the visual comparison
of empirical and theoretical cumulative distributions for various distributions fitted
on a same data set. Function \texttt{cdfcomp} must be called with a first argument 
corresponding to a list of objects of class \texttt{fitdist}, and optionnaly
further arguments to customize the plot, as in the following example comparing the fit of
Weibull, lognormal and gamma distributions to ``groundbeef'' data set.
<<echo=TRUE, fig=FALSE>>=
    fg <- fitdist(groundbeef$serving,"gamma")
    fln <- fitdist(groundbeef$serving,"lnorm")
    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
    xlab="serving sizes (g)",lwd=2)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
    xlab="serving sizes (g)",lwd=2)
@
  \caption{Comparison of CDF plots of various distributions fitted on continuous data}
  \label{compcdf}
\end{figure}

@@

<<echo=TRUE, fig=FALSE>>=
plot(fnb)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(fnb)
@
  \caption{Plot of the fit of a discrete distribution}
  \label{plotdiscfit}
\end{figure}

@@

% ADD the definition of these statistics !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
EQUATIONS TO BE ADDED

As giving  more weight to distribution tails, 
Anderson-Darling statistics is of special interest where it is important to place equal 
emphasis on fitting a distribution at the tails as well as the main body, 
as it is often the case in risk assessment \cite{Cullen99,Vose10}.

% ADD AN EXAMPLE with comparison !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
Nevertheless, this statistics should 
be used cautiously when comparing fits of various distributions, keeping in mind that the weighting of each
cdf quadratic difference is dependent of the theoretical distribution.

@@

Even if specifically recommended for discrete distributions, the Chi-squares statistic may also be used for
continuous distributions (see~\cite{fitdistrplus} for examples).

@@

Goodness-of-fit tests may be used carefully. As for any null-hypothesis significance
test, the non reject of the null hypothesis dose not imply its acceptation. However, this
misinterpretation of p-values is very common and comes from the wrong assumption that
absence of evidence is evidence of absence~\cite{Altman96}.
On the contrary, in some cases, especially on very big datasets, 
even if the null hypothesis is rejected a fitted distribution may be 
chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
plots do not show strong differences between empirical and theoretical distributions.

@@

Using censored data such as those coded in the ``smokedfish'' data set,
the empirical distribution may be plotted using the function \texttt{plotdistcens}.
Data are reported directly as segments for interval, left and right censored data, 
and as points for non-censored data. Before plotting, observations are ordered and a rank r
is associated to each of them. Left censored observations are ordered
first, by their right bounds. Interval censored and non censored observations
are then ordered by their mid-points and, at last, right censored observations are
ordered by their left bounds. If argument \texttt{leftNA} (resp. \texttt{rightNA}) is finite,
left censored (resp. right censored) observations are considered as interval censored
observations and ordered by mid-points with non-censored and interval censored data.
It is sometimes necessary to fix \texttt{leftNA} or \texttt{rightNA} to a realistic 
extreme value, even if not exactly known, to obtain a reasonable global ranking of 
observations. After ranking, each of the n observations is plotted as a point (one x-value) 
or a segment (an interval of possible x-values),
with an y-value equal to r/n, r being the rank of each observation in the global ordering
previously described.

Let us see the sulting plot for ``smokedfish'' data set after classical transformation
of microbila counts in decimal logarithm.

<<echo=TRUE, fig=FALSE>>=
log10C <- data.frame(left=log10(smokedfish$left),right=log10(smokedfish$right))
plotdistcens(log10C)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plotdistcens(log10C)
@
  \caption{CDF plot of a censored data set as}
  \label{plotdistcens}
\end{figure}

@@

As for non censored data, one or more parametric distributions may then 
be fitted to the censored data set, one at a time, but using in this case 
Function \texttt{fitdistcens}. This function estimates the distribution parameters
by maximizing the likelihood given by following equation for censored data.
As \texttt{fitdist}, it returns the results of the fit of any parametric distribution 
to a data set as an S3 class object that may be easily printed, 
summarized or plotted.
% ADD the equation !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
EQUATION TO ADD

@@

\subsection{Goodness-of-fit plot}
\label{censored GOF graph}
%%%%%%%%%%
Only one goodness-of-fit plot is provided for censored data, correponding to the theoretical 
cumulative distribution function added to the plot of censored data presented
in Section~\ref{censored graph}.

<<echo=TRUE, fig=FALSE>>=
plot(flog10C)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(flog10C)
@
  \caption{Goodness-of-fit CDF plot for a fit of a continuous distribution
  on censored data}
  \label{plotfitdistcens}
\end{figure}

@@

%Computational forms of these distances are given in Table~\ref{MGEdist}.
% ADD TABLE with formulas !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
TABLE with formulas TO BE ADDED

@@

Maximum goddness-of-fit estimation may also be useful to give
more weight to data at one tail of the distribution. In ecotoxicology, species sensitivity distributions such 
as those presented in \cite{Hose04} are often
fitted by a lognormal distribution (or another parametric distribution)
so as to estimate a low percentile, often 5$\%$ percentile,
named the hazardous concentration 5$\%$ (HC5). This value is then interpreted as a value of the
contaminant concentration protecting 95$\%$ of the species. In this context, one may consider
to fit the parametric distribution by giving more weight to the left tail of the empirical distribution
such as in the following example using left tail Anderson-Darling distances of first or second order.

>>>>

Fitting distributions to data is a very common task in statistics. It consists in choosing
a probability distribution that gives a good representation of a statistical
variable. It requires judgment and expertise and generally needs an iterative process of
distribution choice, parameter estimation, and quality of fit evaluation. 
Function \texttt{fitdistr} in the R package \texttt{MASS} \cite{MASS} 
is a well known general-purpose maximum-likelihood
fitting routine for the parameter estimation step in R. Other steps of
the process may be developed using R \cite{Ricci05}. Our first objective by developping
package \texttt{fitdistrplus} \cite{fitdistrplus} 
was to provide R users a set of functions dedicated to help the overall
process of fitting a univariate parametric distribution to data.

@@

% A changer en cas de mise a jour de fitdistrplus !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
This manuscript reviews the various features of version 0.3-4 of \texttt{fitdistrplus}. The package
is available from the Comprehensive R Archive Network at \url{http://cran.r-project.org/package=fitdistrplus}.
The development version of the package is located at R-forge as one the packages of the 
project ``Risk Assessment with R''
(\url{http://r-forge.r-project.org/projects/riskassessment/})
The following command will load the package.
<<results=hide>>=
library(fitdistrplus)
@

\subsection{Running examples}
\label{Examples}
%%%%%%%%%%%
For illustrating the use of various functions of package \texttt{fitdistrplus}, we will use
four examples published in various biological areas, 
corresponding to data sets included in the package.

@@

The ``Toxocara'' data set corresponds to the observation of a discrete variable, 
the number of \emph{Toxocara cati} parasites present in digestive tract,
on a random sample of feral 
cats living on Kerguelen island (\cite{Fromont01}).
<<echo=TRUE>>=
data(toxocara)
str(toxocara)
@

The ``smoked fish'' data set corresponds to the observation of a continuous censored variable, 
the \emph{Listeria monocytogenes} microbial concentration, on a random
sample of smoked fish distributed on the Belgian market in the period
2005 to 2007 (\cite{Busschaert10}).
Censored data are coded within 2 columns named left and right, describing
each observed value of \emph{Listeria monocytogenes} concentration 
(in $CFU.g^{-1}$) as an interval. 
The left column contains either \texttt{NA} for
left censored observations, the left bound of the interval for interval censored
observations, or the observed value for non-censored observations. The right
column contains either \texttt{NA} for right censored observations, the right bound of
the interval for interval censored observations, or the observed value for noncensored
observations.
<<echo=TRUE>>=
data(smokedfish)
str(smokedfish)
@

@@

<<echo=TRUE, fig=FALSE>>=
plotdist(groundbeef$serving)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, fig=TRUE>>=
plotdist(groundbeef$serving)
@
  \caption{Density and cdf plots of an empirical distribution for a continuous variable 
  (serving size from the ``ground beef'' data set)}
  \label{plotdistcont}
\end{figure}

@@

<<echo=TRUE, fig=FALSE>>=
plotdist(toxocara$number,discrete = TRUE)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, fig=TRUE>>=
plotdist(toxocara$number,discrete = TRUE)
@
  \caption{Density and cdf plots of an empirical distribution for a discrete variable
  (number of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
  \label{plotdistdisc}
\end{figure}

@@

\subsection{Empirical basis for selecting candidate distributions}
\label{Empirical}
%%%%%%%%%% 
Descriptives statistics may help the choice of good candidates to describe an empirical distribution
among a family of parametric distributions. Especially the skewness and kurtosis are useful for this
purpose. The concept of skewness relates to deviations from symmetry of the distribution.
The normal distribution has a skewness of zero. A positive (resp. negative) skewness indicates 
that the right (resp. left) tail of the distribution is more extended than the left (resp. right) one.
The concept of kurtosis relates to the tail weight. The normal distribution has a kurtosis of 3.
Distributions with a higher kurtosis are said to be leptokurtic, with heavier tails, such as the 
logistic distribution, while distributions with a smaller kurtosis are said platykurtic, with lighter 
tails, such as the uniform distribution.

Function \texttt{descdist} provides calculations of classical descriptive statistics
(minimum, maximum, median, mean, sample sd) and by default 
unbiased estimations of skewness and 
Pearsons's kurtosis values. Nevertheless estimations of skewness and kurtosis
are unbiased only for normal distributions and estimated values are thus only indicative.
A skewness-kurtosis plot such as the one proposed by \cite{Cullen99} is also provided for the 
empirical distribution (Figure~\ref{Cullenplotcont}). 
On this plot, values for common distributions are displayed as tools 
to help the choice of distributions to fit to data. For some distributions (normal, uniform,
logistic, exponential for example), there is only one possible value for the skewness and the kurtosis
 and the distribution is thus represented by a point on the plot. For other distributions, 
areas of possible values are represented, consisting in lines (as for gamma and lognormal distributions), 
or larger areas (as for beta distribution).

Skewness and kurtosis are known not to be robust. In order to take into account the uncertainty 
of the estimated values of kurtosis and skewness from data, the data set may be boostraped by 
fixing the argument \texttt{boot} to an integer above 10. Values of skewness and kurtosis 
corresponding to bootstrap samples are then computed and reported on the 
skewness-kurtosis plot. Below is a call to function \texttt{descdist} to describe the distribution 
of the serving size from the ``ground beef'' data set and to draw the corresponding skewness-kurtosis
plot (Figure~\ref{Cullenplotcont}). Looking at the results on this example
with a positive skewness and a kurtosis not far from 3,
the fit of three common right-skewed distributions could be considered, Weibull, gamma and 
lognormal distributions.

<<echo=TRUE, fig=FALSE>>=
descdist(groundbeef$serving,boot=1000)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
descdist(groundbeef$serving,boot=1000)
@
  \caption{Skewness-kurtosis plot for a continuous variable 
  (serving size from the ``ground beef'' data set)}
  \label{Cullenplotcont}
\end{figure}

For discrete variables, such as the number of \emph{Toxocara cati} parasites from 
the ``Toxacara'' data set,  skewness and kurtosis values or set of values of Poisson and negative binomial
distributions are represented in the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}), 
together with values for the
normal distribution, to which discrete distributions may converge. 
Looking at the skewness-kurtosis plot (Figure~\ref{Cullenplotdisc}) obtained
for the number of \emph{Toxocara cati} parasites from 
the ``Toxacara'' data set, one could try the fit of Poisson and negative-binomial distributions.

<<echo=TRUE, fig=FALSE>>=
descdist(toxocara$number,discrete = TRUE,boot=1000)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
descdist(toxocara$number,discrete = TRUE,boot=1000)
@
  \caption{Skewness-kurtosis plot for a discrete variable
  (number of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
  \label{Cullenplotdisc}
\end{figure}

@@

\section{Fit of a distribution by maximum likelihood estimation}
\label{FIT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter estimation}
\label{Estim}
%%%%%%%%%%
Once selected, one or more parametric distributions may be fitted to the data set, one at a time, using
Function \texttt{fitdist}. By default, distribution parameters $\theta$ are estimated by maximizing the
likelihood defined as:
  \begin{equation}
    \label{likelihood}
    L(\theta)=\prod_{i=1}^n f(y_{i}|\theta)
  \end{equation}
with $y_{i}$ the $n$ observations of variable $y$ and $f$ the density function of the fitted 
parametric distribution.

% TO DO : change the formalism of this equation

@@

Function \texttt{fitdist} returns the results of the fit of any parametric distribution 
to a data set as an S3 class object that may be easily printed, 
summarized or plotted (see Figure~\ref{plotcontfit} in Section~\ref{GOF plots}). 
The parametric distribution must be a classically defined R distributions,
with at least d, p and q functions respectively for the density cdf and quantile functions
(for example \texttt{dnorm}, \texttt{pnorm} and \texttt{qnorm} for the normal distribution).
The name of the fitted distribution is specified in the first argument by its classical 
abbreviation used as the second part of d, p and q functions (for example ``norm''
for the normal distribution. Numerical results returned by Function \texttt{fitdist} are 
parameter estimates with estimated standard errors computed from the estimate
of the Hessian matrix at the maximum likelihood solution, correlation matrix
between parameter estimates, the loglikelihood, the Akaike and the Schwarz information criteria 
(so called AIC and BIC).
Below is a call to function \texttt{fitdist} to fit a Weibull distribution 
to the serving size in the ``ground beef'' data set.
<<echo = TRUE,fig = FALSE>>=
fw <- fitdist(groundbeef$serving,"weibull")
print(fw)
summary(fw)
@

@@

\subsection{Goodness-of-fit plots}
\label{GOF plots}
%%%%%%%%%%
The plot of an object of class \texttt{fitdist} provides two types of results depending of the nature 
of the distribution, continuous or discrete. For continuous distributions, 
four goodness-of-fit plots are provided : a draw of pdf curve and histogram together, an cdf plot
of both empirical and theoretical distributions, a Q-Q plot
(plot of the quantiles of the theoretical fitted distribution (x-axis) against the empirical quantiles of the data)
and a P-P plot (i.e. for each value of the data set, plot of the cumulative density function of the fitted distribution
(x-axis) against the empirical cumulative density function (y-axis)) are also given (\cite{Cullen99}).
The Q-Q plot emphasizes the lack-of-fit at the distribution tails while the P-P plot emphasizes
the lack-of-fit at the distribution center.
As an example, let us look at the plot of the previous fit of a Weibull distribution to the``groundbeef'' data set
(Figure~\ref{plotcontfit}). The fit is not perfect,
especially in the center of the distribution, but seems correct while looking at the tails.

<<echo=TRUE, fig=FALSE>>=
plot(fw)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(fw)
@
  \caption{Plot of the fit of a continuous distribution (a Weibull
  distribution fitted to serving sizes from the ``ground beef'' data set)}
  \label{plotcontfit}
\end{figure}

For continuous distributions, Function \texttt{cdfcomp} enables the visual comparison
of empirical and theoretical cumulative distributions for various distributions fitted
on a same data set. Function \texttt{cdfcomp} must be called with a first argument 
corresponding to a list of objects of class \texttt{fitdist}, and optionaly
further arguments to customize the plot, as in the following example comparing the fit of
Weibull, lognormal and gamma distributions to ``groundbeef'' data set (Figure~\ref{compcdf}).
<<echo=TRUE, fig=FALSE>>=
    fg <- fitdist(groundbeef$serving,"gamma")
    fln <- fitdist(groundbeef$serving,"lnorm")
    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
    xlab="serving sizes (g)",lwd=2)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
    cdfcomp(list(fw,fln,fg),legendtext=c("Weibull","lognormal","gamma"),
    xlab="serving sizes (g)",lwd=2)
@
  \caption{Comparison of CDF plots of various distributions fitted on continuous data
  (Weibull, gamma and lognormal
  distributions fitted to serving sizes from the ``ground beef'' data set)}
  \label{compcdf}
\end{figure}

@@

<<echo=TRUE, fig=FALSE>>=
plot(fnb)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(fnb)
@
  \caption{Plot of the fit of a discrete distribution (a negative binomial distribution
  fitted to numbers of \emph{Toxocara cati} parasites from the ``Toxocara'' data set)}
  \label{plotdiscfit}
\end{figure}

@@

% ADD the definition of these statistics !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
%EQUATIONS TO BE ADDED

As giving  more weight to distribution tails, 
Anderson-Darling statistics is of special interest where it is important to place equal 
emphasis on fitting a distribution at the tails as well as the main body, 
as it is often the case in risk assessment \cite{Cullen99,Vose10}. 
% ADD AN EXAMPLE with comparison !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
Nevertheless, this statistics should 
be used cautiously when comparing fits of various distributions, keeping in mind that the weighting of each
cdf quadratic difference is dependent of the theoretical distribution.

@@

Even if specifically recommended for discrete distributions, the Chi-squared statistic may also be used for
continuous distributions (see~\cite{fitdistrplus} for examples).

@@

Goodness-of-fit tests may be used carefully. As for any null-hypothesis significance
test, the non reject of the null hypothesis dose not imply its acceptation. However, this
misinterpretation of p-values is very common and comes from the wrong assumption that
absence of evidence is evidence of absence ~\cite{Altman96}.
On the contrary, in some cases, especially on very big datasets, 
even if the null hypothesis is rejected, a fitted distribution may be 
chosen as the best one among simple distributions to describe an empirical distribution, if the goodness-of-fit
plots do not show strong differences between empirical and theoretical distributions.

@@

Using censored data such as those coded in the ``smokedfish'' data set,
the empirical distribution may be plotted using the function \texttt{plotdistcens}.
By default this function uses the EM approach of Turnbull \cite{Turnbull74} to compute the overall
empirical cdf curve with confidence intervals, by calls to functions \texttt{survfit}
and \texttt{plot.survfit} from the \texttt{survival} package.
Let us see such a plot for ``smokedfish'' data set after classical transformation
of microbial counts in decimal logarithm (Figure~\ref{plotdistcens}).

% ------------------ previous method -----------------------------
%Data are reported directly as segments for interval, left and right censored data, 
%and as points for non-censored data. Before plotting, observations are ordered and a rank r
%is associated to each of them. Left censored observations are ordered
%first, by their right bounds. Interval censored and non censored observations
%are then ordered by their mid-points and, at last, right censored observations are
%ordered by their left bounds. If argument \texttt{leftNA} (resp. \texttt{rightNA}) is finite,
%left censored (resp. right censored) observations are considered as interval censored
%observations and ordered by mid-points with non-censored and interval censored data.
%It is sometimes necessary to fix \texttt{leftNA} or \texttt{rightNA} to a realistic 
%extreme value, even if not exactly known, to obtain a reasonable global ranking of 
%observations. After ranking, each of the n observations is plotted as a point (one x-value) 
%or a segment (an interval of possible x-values),
%with an y-value equal to r/n, r being the rank of each observation in the global ordering
%previously described.

<<echo=TRUE, fig=FALSE>>=
log10C <- data.frame(left=log10(smokedfish$left),right=log10(smokedfish$right))
plotdistcens(log10C)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plotdistcens(log10C)
@
  \caption{CDF plot of censored data (microbial counts from 
   the ``smoked fish'' data set)}
  \label{plotdistcens}
\end{figure}

@@

As for non censored data, one or more parametric distributions may then 
be fitted to the censored data set, one at a time, but using in this case 
Function \texttt{fitdistcens}. This function estimates the distribution parameters
by maximizing the likelihood given by following equation for censored data.
As \texttt{fitdist}, it returns the results of the fit of any parametric distribution 
to a data set as an S3 class object that may be easily printed, 
summarized or plotted.
% ADD the equation !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@

\subsection{Goodness-of-fit plot}
\label{censored GOF graph}
%%%%%%%%%%
Only one goodness-of-fit plot is provided for censored data (Figure~\ref{plotfitdistcens}), corresponding to the theoretical 
cumulative distribution function added to the plot of censored data presented
in Section~\ref{censored graph}.

<<echo=TRUE, fig=FALSE>>=
plot(flog10C)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
plot(flog10C)
@
  \caption{Goodness-of-fit CDF plot for a fit of a continuous distribution
  on censored data (a lognormal distribution fitted to microbial counts from 
   the ``smoked fish'' data set)}
  \label{plotfitdistcens}
\end{figure}

@@

%Computational forms of these distances are given in Table~\ref{MGEdist}.
% ADD TABLE with formulas !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@

Maximum goddness-of-fit estimation may also be useful to give
more weight to data at one tail of the distribution. In ecotoxicology, species sensitivity distributions such 
as those presented in \cite{Hose04} are often
fitted by a lognormal distribution (or another parametric distribution)
so as to estimate a low percentile, often 5$\%$ percentile,
named the hazardous concentration 5$\%$ (HC5). This value is then interpreted as a value of the
contaminant concentration protecting 95$\%$ of the species. In this context, one may consider
to fit the parametric distribution by giving more weight to the left tail of the empirical distribution
such as in the following example using left tail Anderson-Darling distances of first or second order
(Figure~\ref{plotfitMGE}).

<<echo=TRUE, fig=FALSE>>=
data(endosulfan)
ATV <-subset(endosulfan,group == "NonArthroInvert")$ATV
flnMGEKS <- fitdist(ATV,"lnorm",method="mge",gof="KS")
flnMGEAD <- fitdist(ATV,"lnorm",method="mge",gof="AD")
flnMGEADL <- fitdist(ATV,"lnorm",method="mge",gof="ADL")
flnMGEAD2L <- fitdist(ATV,"lnorm",method="mge",gof="AD2L")
cdfcomp(list(flnMGEKS,flnMGEAD,flnMGEADL,flnMGEAD2L),
xlogscale = TRUE,main="",
legendtext = c("Kolmogorov-Smirnov (KS)","Anderson-Darling",
"Left-tail Anderson-Darling","Left tailed Anderson-Darling of second order"),cex=0.7,
xlegend = 500, ylegend = 0.15)
@
\begin{figure}[htb]
  \centering
<<echo=FALSE, results=hide, fig=TRUE>>=
cdfcomp(list(flnMGEKS,flnMGEAD,flnMGEADL,flnMGEAD2L),
xlogscale = TRUE,main="",
legendtext = c("Kolmogorov-Smirnov (KS)","Anderson-Darling",
"Left-tail Anderson-Darling","Left tailed Anderson-Darling of second order"),cex=0.7,
xlegend = 500, ylegend = 0.15)
@
  \caption{Comparison of one distribution fitted by maximum goodness-of-fit using
  various goodness-of-fit distances (a lognormal distribution fitted to acute toxicity values
  from the ``endosulfan'' data set)}
  \label{plotfitMGE}
\end{figure}
